---
title: "Project 3 Skill Analysis"
author: "Donald Butler"
date: "10/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

  
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(tidytext)
library(janeaustenr)
```

```{r}

urlfile<-"https://raw.githubusercontent.com/nolivercuny/data607-team-6-project-3/master/data/job_listings_final.csv" 
jobdat <- read_csv(url(urlfile))

#view short file summary and class
jobdat<-data_frame(jobdat)
glimpse(jobdat)

```

Created ngrams of 1, 2, and 3 words, filtered out the common stop_words, and then did a count of each and filtered down to the rows that occured at least 10 times within our description dataset. I then exported the results which contained 6015 rows to a file which I loaded into Excel to determine which were relevant job skills and which were not. 

```{r text}

jobdat_1gram <- jobdat %>%
  unnest_tokens(ngram,description,token='ngrams',n=1,format='text',drop=TRUE,to_lower=TRUE) %>%
  filter(!ngram %in% stop_words$word) %>%
  count(ngram,sort = TRUE) %>%
  filter(n >= 10)

jobdat_2gram <- jobdat %>%
  unnest_tokens(ngram,description,token='ngrams',n=2,format='text',drop=TRUE,to_lower=TRUE) %>%
  separate(ngram,c('word1','word2'),sep = " ") %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) %>%
  unite(ngram, c('word1','word2'), sep = " ") %>%
  count(ngram,sort = TRUE) %>%
  filter(n >= 10)

jobdat_3gram <- jobdat %>%
  unnest_tokens(ngram,description,token='ngrams',n=3,format='text',drop=TRUE,to_lower=TRUE) %>%
  separate(ngram,c('word1','word2','word3'),sep = " ") %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) %>%
  filter(!word3 %in% stop_words$word) %>%
  unite(ngram, c('word1','word2','word3'), sep = " ") %>%
  count(ngram,sort = TRUE) %>%
  filter(n >= 10)

jobdat_ngrams <- jobdat_1gram %>%
  rbind(jobdat_2gram) %>%
  rbind(jobdat_3gram) %>%
  arrange(desc(n))

head(jobdat_ngrams,25)

jobdat_ngrams %>%
  write.table(file = './jobdat_ngrams.csv',quote = FALSE, sep = '\t', row.names = FALSE)

```

I generated a file of applicable skills, along with common alternative spellings, ML & Machine Learning. On lines with multiple alternatives, they are separated by a pipe |, for RegEx comparison.

```{r}

(jobskills <- read_csv('https://raw.githubusercontent.com/dab31415/DATA607/main/Projects/Project_3/JobSkills.csv'))

```

I loop though the list of skills and use regex to determine if the skill is listed within the description text. A new attribute is created in the jobs data frame that indicates if the skill is required.

```{r}

for (i in 1:nrow(jobskills)) {
  jobdat[,ncol(jobdat) + 1] <- str_detect(jobdat$description, regex(paste('[^A-Z0-9]',jobskills[i,1],'[^A-Z0-9]',sep = ''),ignore_case = TRUE))
  colnames(jobdat)[ncol(jobdat)] <- as.character(jobskills[i,1])
}

glimpse(jobdat)

```