---
title: "Project 3"
author: "Nick Oliver"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
editor_options: 
  chunk_output_type: console
---

# Project 3 Analysis

## Setup

### Load Libraries


```{r results=FALSE, message=FALSE, warning=FALSE}
library(DBI)
library(tidyverse)
library(ggplot2)
library(kableExtra)
```

### Open DB connection

```{r}
dbLocation <-"/Users/t0pth4t/Downloads/CUNY/DATA607/data607-team-6-project-3/data/project3_job_skills.db"
print(dbLocation)
con <- dbConnect(RSQLite::SQLite(), dbname = dbLocation)
```

## Which companies require the most skills?

- Select job listings grouped by company and count the number of skills required. 
- Then select number of job listings by company. 
- Join the two dataframes and add a column which is the mean number of skills required by taking the total number of required skills divided by the total number of listings
- Then arrage by higest number of skills required, select to 10, and put in a table.

```{r}
companySkillsCountDf <- dbGetQuery(con, "select c.company_name, count(jls.skill) as required_skills
from job_listing jl
join job_listing_skill jls on jl.id = jls.job_listing_id
join company c on c.id = jl.company_id
group by company_name
order by required_skills desc;
")

companyJobListingsCountDf <- dbGetQuery(con, "
select c.company_name, count(jl.id) as number_job_listings
from job_listing jl
join company c on c.id = jl.company_id
group by company_name
order by number_job_listings desc; ")

joinedDf <- left_join(companyJobListingsCountDf, companySkillsCountDf, by="company_name") %>%
  mutate(mean_skills_per_listing = round(required_skills / number_job_listings))


joinedDf %>% 
  select(company_name, mean_skills_per_listing) %>% 
  arrange(desc(mean_skills_per_listing)) %>%
  top_n(10, mean_skills_per_listing) %>%
  kable(caption = "Companies Requiring Most Skills", format = "html", col.names = c("Company Name", "Mean Number Skills Required")) %>% 
  kable_styling("striped", "hover") %>% 
  scroll_box(width = "100%")
```

### Which companies require fewest skills?

- Take the same dataframe as above except arrange in reverse order and take the bottom 10 companies

```{r}
joinedDf %>% 
  select(company_name, mean_skills_per_listing) %>% 
  arrange(mean_skills_per_listing) %>%
  top_n(-10,mean_skills_per_listing) %>%
  kable(caption = "Companies Requiring Fewest Skills", format = "html", col.names = c("Company Name", "Mean Number Skills Required")) %>% 
  kable_styling("striped", "hover") %>% 
  scroll_box(width = "100%")
```


## What Companies Do Not Require The Most Popular Skill?

```{r}
topSkillDf <- dbGetQuery(con,"select count(1), skill
from job_listing_skill
group by skill
order by count(1) desc
limit 1;")

print(topSkillDf$skill)

allListingsDf <- dbGetQuery(con, "select jl.id, jl.job_title, c.company_name, jls.skill, jl.salary
from job_listing jl
join company c on c.id = jl.company_id
join job_listing_skill jls on jl.id = jls.job_listing_id;")

groupedSkillsDf <- allListingsDf %>% group_by(id, company_name, job_title, salary) %>% summarise(skills = paste(skill, collapse = ","))


noPythonJobsDf <- groupedSkillsDf %>% filter(!grepl("Python", skills))

print(scales::percent(nrow(noPythonJobsDf) / nrow(groupedSkillsDf)))

```

**Roughly 28% of the jobs do not require python.**

## Teardown

### Close DB Connection

```{r}
dbDisconnect(con)
```



